Towards a classification framework for social machines

### 2013-04-04 latex/13socm07-shadbolt.tex
Updated to version submitted to sheridan

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

\usepackage{url}

@@

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Nigel Shadbolt,^{1} Daniel Alexander Smith,^{2} Elena Simperl,^{3} Max Van Kleek,^{4} Yang Yang,^{5} Wendy Hall^{6}
       \affaddr{Web and Internet Science Group}\\
       \affaddr{University of Southampton}\\
       \affaddr{Southampton, UK}\\
       \email{\{^{1}nrs, ^{2}ds, ^{3}es, ^{4}emax, ^{5}yy, ^{6}wh@ecs.soton.ac.uk}
}

% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

The state of the art in human interaction with computational systems blurs the line between
computations performed by machine logic and algorithms, and those that result from input by
humans, arising from their own psychological processes and life experience. Current
socio-technical systems, known as `social machines' exploit the large-scale interaction of
humans with machines. Interactions that are motivated by numerous goals and purposes
including financial gain, charitable aid, and simply for fun. In this paper we explore
the landscape of social machines, both past and present, with the aim of defining an
initial classificatory framework. Through a number of knowledge elicitation
and refinement exercises we have identified the polyarchical relationship between
infrastructure, social machines, and large-scale social initiatives. Our initial framework
describes classification constructs in the areas of {\it contributions}, {\it participants}, and
{\it motivation}. We present an initial characterization of some of the most popular social machines, as
an initial validation and demonstration of the use of the identified constructs. We believe that it is important to undertake an analysis of the behaviour and phenomenology of social machines, and of their growth and evolution over time. Our future work will seek to elicit additional opinions, classifications and validation from a wider audience, to produce a comprehensive framework for the description, analysis and comparison of social machines.

@@

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

%\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Social machines: an introduction}
Once upon a time `machines' were programmed by programmers and used by users. The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites. Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them and treat each such compound as a `social machine' --- a machine in which the two aspects are seamlessly interwoven. This was the insight behind Berners-Lee's original characterisation of the concept of a social machine \cite{bernerslee1999}. This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.

Essentially social machines can be characterised as assemblies of manually executed and machine-driven (as in `automatised') services and the interaction of such services. A traditional database, with users looking up records independently of each other and an administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship. When we fill out a form (e.g., health information, birds spotted in a garden, new construction sites on the way to work) the systems supporting this activity are minimal forms of social machines because the users are part of the `social computation', in this case the data creation and collection process facilitated through the site. This social component becomes richer when the database is curated by members of the broader community (e.g., Wikipedia) and when the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through individual or joint activities of the participants.

In general we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand. However, as technology becomes more and more ubiquitous, many of the challenges we face will increasingly require solutions that rely on both of these axes: a sophisticated combination of data-intensive, complex automation, and deep community involvement. This suggests the need for new types of systems to tackle these emerging challenges, and these systems will not be able to be built and used in a sustainable way without a thorough understanding of the science and engineering of (the continuum of) social machines.

@@

The current version of our framework is the result of a knowledge elicitation process performed internally in our lab. Our goal is to present it to the broader community and then to extend and revise it based on feedback in order to achieve a shared understanding of the basic notions and of the associated terminology. We aim to develop a descriptive framework to enable a systematic study of existing and future social machines. We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of such systems. Using a common conceptualisation allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored. They are given a means to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation. For developers and operators, the framework may inform the engineering of new systems in terms of critical features and community development.

@@

The remaining sections are organised as follows. In order to pin down what constitutes a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} provides a summary of this comparative analysis. In Section \ref{sec:framework} we present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via repertory grid analysis. We conclude with our proposed future work, including details concerning the evaluation of the framework and reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed framework for the field.

\section{Comparison with related fields}
\label{sec:comparison}
Turing machines were once conceived as logical automata. We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one according to which machines facilitate a wide range of human interactions. The emergence of `Computer-supported Collaborative Work (CSCW)' \cite{grudin1994computer} is representative of the early phases of this trend. This initial concept has evolved to become the broader field of `Computer-supported collaboration' (CSC).

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and the extent of mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society. This trend towards `prosumerism' is finding more and more adopters in the public and private sector. Governments and enterprises are not only becoming active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, service offerings, and policy formulation. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'. As we argue below, these terms are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/socialmachinesscope_new.png}
\caption{Social machines and related areas} \label{fig:socialmachine}
\end{center}
\end{figure}

Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account the information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales inconceivable in the past, and to involve highly diverse and geographically distributed participants. A similar concept, though broader scoped, is collective intelligence, defined in \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is with regard to the extent and role of automation. In a social machine, human and computational intelligence coalesce in order to achieve a given purpose. While wisdom of the crowds and collective intelligence place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems in which the two types of components co-exist. As such, theoretical and empirical insights from these two related areas are useful to understand the dynamics of the social structures underlying a social machine, but they are definitely not the only ingredient needed to build operational systems. The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering by which a system is developed, tested, and updated are equally important.

One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}. At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, whilst at the same time ensuring in-time access to a potentially infinite pool of skills and resources not available before the advent of social computing technologies. The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}. Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding. These kinds of task are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete. The key difference between social machines and open innovation and crowdsourcing is at the level of the interaction between the social and the machine-driven processing components. The novelty these two other approaches bring in lies in their use of a much larger pool of human resources than traditional work environments \cite{quinn2011human}. There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution. These aspects are equally important for social machines, which, however, also look into the principled combination of human and computational capabilities and the technical means to support them. In comparison to human computation, social machines cover a wider range of scenarios. Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.

An analogous line of reasoning illustrates the overlap between social machines and social computing \cite{parameswaran2007research}. The latter is an area of computer science that refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}. As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort. This distinction is even stronger in the case of social machines, which regard the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.

@@

In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate a set of elements -- for example, prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct --- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two system. The user then rates every element with value from $1$ to $5$ on this construct, where $1$ represents an element that is purely `For Science', and $5$ one that mainly serves `socialising' purposes. In triadic presentation the repertory grid software lists three elements and asks the user to generate a construct where one of the elements is different to the other two. We asked computer science researchers familiar with the field to create their own repertory grids, generating elements from their own knowledge, and creating constructs using this technique. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of a social machine and their most distinctive properties, we allowed the participants to choose the systems they are familiar with and describe them in their own terms.

@@

While determining the intersection of elements was straightforward, the consolidation of the
constructs required a more thorough process. We manually grouped the constructs into rough clusters, based around the areas they cover. We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to reduce redundancies and overlap. Our aim was to cut the classification space down to a manageable size, while ensuring that all constructs that were initially elicited were represented in the final set. This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organised in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives. In our analysis we did not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users. It was apparent that the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavours) in the grids provided by the ten experts, others such as the different types of workflows synthesising human and automatised capabilities of a social machine were less common. As discussed in Section \ref{sec:comparison} we believe that the interaction between these two types of services is an essential part of the theory and practice of social machines, being one of the elements that distinguishes them from related areas which are biased towards one or the other. Extending the framework with constructs capturing these aspects (see, for instance, \cite{quinn2011human} for examples in the context of human computation) is part of our future work.

@@

\subsection{Constructs}
The clustering process described in Section \ref{sec:methodology} yielded both a
reduction in the overall number of constructs (in particular an
elimination of over-represented areas) and the identification of three
themes: first the nature of tasks, problems and activities
performed through and using the machine, second the users and roles within
each machine, and third, motivation for participation.

@@

\section{Future work: building a social machine around the classification framework}
In this paper we presented a first outline of a description space for social machines. We introduced the key features of social machines, and explained how this concept differs from related areas such as collective intelligence, crowdsourcing, and human computing. We then discussed the basic constructs of a classification framework created through the repertory grid elicitation technique. In its current version the framework consists of $31$ constructs clustered according to the main components of social machines: social and machine-driven services, and the interactions between them. Our future work will seek to revise this initial list of constructs. Most notably, we would like to add a number of constructs to emphasize the importance of the interaction aspect in the theory, engineering, and operation of social machines. This includes constructs related to the human-computer workflows supported (e.g., human intelligence replacing automatic algorithms as in GWAPs vs the different bots available on Wikipedia complementing collaborative manual editing), the way inputs from the crowd are validated, aggregated and turned into a useful outcome, as well as issues related collaboration and coordination, and interface and communication design. In addition, we plan to engage with the broader researcher community by setting up a social machine to facilitate discussions and feedback collection. One specific idea that is currently under development is to build a microtask environment, including specific game elements, in which participants are asked to provide answers to atomic challenges that rate and compare a pair of social machine instances according to a construct in our framework. This classificatory work is part of a larger project in which we seek to understand the theory and practice of building social machines and where we might in future anticipate what makes social machines scale or fail.

>>>>

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ tob
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{www13-companion-accepted}

\clubpenalty=10000 
\widowpenalty = 10000

\usepackage{ifpdf}
\ifpdf
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\else
\fi
\usepackage{url}

@@

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Nigel R. Shadbolt, Daniel A. Smith, Elena Simperl, Max Van Kleek, Yang Yang, Wendy Hall\\
       \affaddr{Web and Internet Science Group}\\
       \affaddr{University of Southampton, UK}\\
       \email{\{nrs,ds,es6g12,emax,yy1402,wh\}@ecs.soton.ac.uk}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

The state of the art in human interaction with computational systems blurs the line between computations performed by machine logic and algorithms, and those that result from input by humans, arising from their own psychological processes and life experience. Current socio-technical systems, known as `social machines' exploit the large-scale interaction of humans with machines. Interactions that are motivated by numerous goals and purposes including financial gain, charitable aid, and simply for fun. In this paper we explore the landscape of social machines, both past and present, with the aim of defining an initial classificatory framework. Through a number of knowledge elicitation and refinement exercises we have identified the polyarchical relationship between infrastructure, social machines, and large-scale social initiatives. Our initial framework describes classification constructs in the areas of {\it contributions}, {\it participants}, and {\it motivation}. We present an initial characterisation of some of the most popular social machines, as demonstration of the use of the identified constructs. We believe that it is important to undertake an analysis of the behaviour and phenomenology of social machines, and of their growth and evolution over time. Our future work will seek to elicit additional opinions, classifications and validation from a wider audience, to produce a comprehensive framework for the description, analysis and comparison of social machines.

@@

% A category with the (minimum) three required fields
\category{H.1.2}{Information Systems}{Models and Principles}[User/Machine Systems]
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Social Machines, Web Science}

\keywords{Social Machines;  Web Science}

\section{What Are Social Machines}
Once upon a time `machines' were programmed by programmers and used by users. The success of the Web has changed this relationship: we now see configurations of people interacting with content and with each other, typified by social Web sites. Rather than drawing a line through such Web-based systems to separate the human and digital parts (as computer science has traditionally done), we can now draw a line around them and treat each such compound as a `social machine' --- a machine in which the two aspects are seamlessly interwoven. This was the insight behind Berners-Lee's original characterisation of the concept of a social machine \cite{bernerslee1999}. This crucial transition in thinking acknowledges the reality of current socio-technical systems, and is essential to underpin any understanding of the science and engineering of their future development towards pervasive ecosystems of co-evolving social machines.

Essentially social machines can be characterised as assemblies of manually executed and machine-driven (as in `automatised') services and the interaction of such services. A traditional database, with users looking up records independently of each other and an administrator responsible for the management of the content, has some of the right ingredients, but there is really no social element in this strict provider-consumer relationship. When we fill out a form (e.g., health information, birds spotted in a garden, new construction sites on the way to work) the systems supporting this activity are minimal forms of social machines because the users are part of the `social computation', in this case the data creation and collection process facilitated through the site. This social component becomes richer when the database is curated by members of the broader community (e.g., Wikipedia) and when the social network adds value implicitly (e.g., Amazon) or explicitly (e.g., Facebook) to the overall system through the individual or joint activities of the participants.

In general we still see a divide between conventional IT systems dedicated to data- and computation-intensive tasks, and Web $2.0$ sites offering some combination of well-known participatory features, in which user-generated content and the underlying social network evolve dynamically and hand-in-hand. However, as technology becomes more and more ubiquitous, many of the challenges we face will increasingly require solutions that rely on both of these axes: a sophisticated combination of data-intensive, complex automation and deep community involvement. This suggests the need for new types of systems to tackle these emerging challenges, and these systems will not be able to be built and used in a sustainable way without a thorough understanding of the science and engineering of (the continuum of) social machines.

@@

The current version of our framework is the result of a knowledge elicitation process performed internally in our lab. Our goal is to present it to the broader community and then to extend and revise it based on feedback in order to achieve a shared understanding of the basic notions and of the associated terminology. We aim to develop a descriptive framework to enable a systematic study of existing and future social machines. We anticipate that the framework will be a useful tool for both researchers in social and computer sciences, and for developers and operators of social machines. Using a common conceptualisation allows the former to become familiar with the landscape, and identify topics of research that so far have remained unexplored or under-explored. They are given a means to observe the effects of specific technical properties of a system configuration on social behaviour, discover design and evolution patterns of such systems, learn about social network formation and dynamics, and devise incentive mechanisms to encourage wide participation. For developers and operators, the framework may inform the engineering of new systems in terms of critical features and community development.

@@

The remaining sections are organised as follows. In order to pin down what constitutes a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} provides a summary of this comparative analysis. In Section \ref{sec:framework} we present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via repertory grid analysis. We conclude with our proposed future work, including details concerning the evaluation of the framework. Finally we reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed classification framework for the field.

\section{Comparison with related fields}
\label{sec:comparison}
Turing machines were once conceived as logical automata. We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one where machines facilitate a wide range of human interactions. The emergence of `Computer-supported Collaborative Work (CSCW)' \cite{grudin1994computer} is representative of the early phases of this trend. This initial concept has evolved to become the broader field of `Computer-supported collaboration' (CSC).

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and the extent of mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopaedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society. This trend towards `prosumerism' is finding more and more adopters in the public and private sector. Governments and enterprises are not only becoming active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, service offerings, and policy formulation. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'. As we argue below they are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/Socialmachine1.png}
\caption{Social machines and related areas} \label{fig:socialmachine}
\end{center}
\end{figure}

Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account the information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales inconceivable in the past, and to involve highly diverse and geographically distributed participants. A similar concept, though broader scoped, is collective intelligence, defined in \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is with regard to the extent and role of automation. In a social machine, human and computational intelligence coalesce in order to achieve a given purpose. Wisdom of the crowds (WoC) and collective intelligence (CI) clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems where the two types of components co-exist. As such, theoretical and empirical insights from WoC and CI are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines. The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering by which a system is developed, tested, and updated are equally important.

One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}. At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, whilst at the same time ensuring in-time access to a potentially infinite pool of skills and resources not available before the advent of social computing technologies. The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}. Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding. These sort of tasks are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete. The key difference between social machines and open innovation and crowdsourcing is at the level of the interaction between the social and the machine-driven processing components. The novelty these two other approaches bring in lies in their use of a much larger pool of human resources then traditional work environments \cite{quinn2011human}. There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution. These aspects are equally important for social machines, which, however, also look into the principled combination of human and computational capabilities and the technical means to support them. In comparison to human computation, social machines cover a wider range of scenarios. Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.

An analogous line of reasoning illustrates the overlap between social machines and social computing \cite{parameswaran2007research}. The latter is an area of computer science which refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}. As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort. This distinction is even stronger in the case of social machines, which regard the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.

@@

In repertory grid elicitation a software tool is used to ask users to describe constructs
that differentiate a set of elements -- for example, prompting the user to create a construct that
differentiates between {\it GalaxyZoo}\footnote{In GalaxyZoo people classify galaxies with collective performance as good as professional astronomers, see: \\ \url{http://www.galaxyzoo.org}.} and {\it Facebook}, as two prominent example of social machines our classification framework should apply to. The user describes the opposing poles of the construct --- in this case, the user may decide on `For Science' and `For socialising' in order to capture the core distinction between the two systems. The user then rates every element with a value from $1$ to $5$ on this construct, where $1$ represents an element that is purely
`For Science', and $5$ one that mainly serves `socialising' purposes. In triadic presentation the repertory grid software presents three elements and asks the user to generate a construct where one of the elements is different to the other two. We asked computer science researchers familiar with the field to create their own repertory grids, generating elements from their own knowledge, and creating constructs using this technique. This exercise led to $10$ grids, the union of which comprised a total of $56$ unique elements (social machines)
and $117$ different constructs (classifying factors). As the aim of this initial phase was to understand how people perceive the notion of social machine and their most distinctive properties, we allowed the participants to choose the social machines they are familiar with and describe them in their own terms.

@@

While determining the intersection of elements was straightforward, the consolidation of the
constructs required a more thorough process. We manually grouped the constructs into rough clusters, based around the areas they cover. We examined each construct to determine which were equivalent, and whether we could re-word or subsume existing constructs to reduce redundancies and overlap. Our aim was to cut the classification space down to a manageable size, while ensuring that all constructs that were initially elicited were represented in the final set. This process involved four of the authors discussing the choices, and resulted in a consolidated set of $31$ constructs organised in four different clusters: popularity, tasks and purpose, participants and roles, and motivation and incentives. In our analysis we will not further consider the first cluster, which primarily covered descriptive information about a given system, such as its perceived level of maturity and current number of users. It was apparent that the constructs were not evenly distributed across clusters; whereas some aspects such as roles of participants occurred frequently (in various flavours) in the grids provided by the ten experts, others such as the different types of workflows synthesising human and automatised capabilities of a social machine were less common. As discussed in Section \ref{sec:comparison} we believe that the interaction between these two types of services is an essential part of the theory and practice of social machines, being one of the elements that distinguishes them from related areas which are biased towards one or the other. Extending the framework with constructs capturing these aspects (see, for instance, \cite{quinn2011human} for examples in the context of human computation) is part of our future work.

@@

\vfill\eject
\subsection{Constructs}
The clustering process described in Section \ref{sec:methodology} yielded both a
reduction in the overall number of constructs (in particular an
elimination of over-represented areas) and the identification of three
themes: first the nature of tasks, problems and activities
performed through and using the machine, second sets of constructs pertaining to users and roles within
each machine, third, motivation for participation.

@@

\section{Future work}
In this paper we presented a first outline of a description space for social machines. We introduced the key features of social machines, and explained how this concept differs from related areas such as collective intelligence, crowdsourcing and human computing. We then discussed the basic constructs of a classification framework created through the repertory grid elicitation technique. In its current version the framework consists of $31$ constructs clustered according to the main components of social machines: social and machine-driven services, and the interactions between them. Our future work will seek to revise this initial list of constructs. Most notably, we would like to add a number of constructs to emphasise the importance of the interaction aspect in the theory, engineering, and operation of social machines. This includes constructs related to the human-computer workflows supported (e.g., human intelligence replacing automatic algorithms as in GWAPs vs the different bots available on Wikipedia complementing collaborative manual editing), the way inputs from the crowd are validated, aggregated and turned into a useful outcome, as well as issues related collaboration and coordination, and interface and communication design. In addition, we plan to engage with the broader researcher community by setting up a social machine to facilitate discussions and feedback collection. One specific idea that is currently under development is to build a microtask environment, including specific game elements, in which participants are asked to provide answers to atomic challenges that rate and compare a pair of social machine instances according to a construct in our framework. This classificatory work is part of a larger project in which we seek to understand the theory and practice of building social machines and where we might in future anticipate what makes social machines scale or fail.

\section{Acknowledgments}
This work is supported under SOCIAM: The Theory and Practice of Social Machines. The SOCIAM Project is funded by the UK Engineering and Physical Sciences Research Council (EPSRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh. The authors would like to thank the entire SOCIAM team for their insightful feedback and lively discussions.
